| Base de datos de abejas Ibéricas \n| Iberian Bees database\n

### 2022-03-12 Manuscript/manuscript.Rmd
manuscript now runs with the current structure

---
title: "| Base de datos de abejas Ibéricas \n| Iberian Bees database\n"

author: Ignasi Bartomeus^1^, Jose B. Lanuza, Thomas Woods^2^, Luisa Carvalheiro^3^, and a lot of coauthors^4^

output:
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7.5
    highlight: null
    reference_docx: Ecosistemas_template.docx

csl: ecosistemas.csl
bibliography: 
  - references.bib
  - knitcitations.bib
---

```{r eval=FALSE, include=FALSE}
#to access the template I ran: 
library(devtools)
install_github("Pakillo/rmdTemplates")

# 2000 words 1-3 figures.
```

```{r include=FALSE}
data <- read.csv("../data/data_clean.csv", stringsAsFactors = FALSE)
master <- read.csv("../data/Iberian_species_masterlist.csv", stringsAsFactors = FALSE)
removed <- read.csv("../data/removed.csv", stringsAsFactors = FALSE)
master$accepted_name <- paste(master$Genus, master$Species, sep = " ")
geo <- length(which(!is.na(data$Latitude) & !is.na(data$Longitude)))
flw <- length(which(!is.na(data$Flowers.visited)))
```

@@

```{r Tabla1, results='asis', echo=FALSE, cache=FALSE}
library("knitr")
metadata <- read.csv("../data/metadata/attributes.csv")
kable(metadata[,-1])
```

@@

```{r Fig2, echo=FALSE, fig.cap="Figura 2. Segundo ejemplo", message=FALSE, warning=FALSE, cache=FALSE}
#code courtesy of Paco Rodriguez.
library(sf)
library(raster)
library(rasterVis)

crs.geo <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
crs.laea <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

dat <- subset(data, !is.na(Latitude) & !is.na(Longitude))
datos.sf <- st_as_sf(dat,
                     coords = c("Longitude", "Latitude"),
                     crs = crs.geo)

@@

ras.counts <- rasterize(datos.laea[, "accepted_name"], ras,
               field = "accepted_name", fun = "count")
#str(datos.laea)
datos.laea$total <- datos.laea$Male + datos.laea$Female + datos.laea$Worker + datos.laea$Not.specified 
ras.counts2 <- rasterize(datos.laea[, "total"], ras,
               field = "total", fun = "sum")
ras2 <- raster(datos.laea, resolution = 30000)
ras.richness <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {length(unique(na.omit(x)))} )
library(vegan)
ras.chao <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {estimateR(t(as.matrix(x = table(x), nrow = 1, ncol = x)))[2,1]} ) #2 chao, 4 ACE #TO DO THIS WELL, we need to recover abundances.

>>>>

---
title: "| Base de datos de abejas Ibéricas \n| Iberian Bees database\n"
author: Ignasi Bartomeus^1^, Jose B. Lanuza, Thomas Woods^2^, Luisa Carvalheiro^3^,
  and a lot of coauthors^4^
output:
  pdf_document: default
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 7.5
    highlight: null
    reference_docx: Ecosistemas_template.docx
csl: ecosistemas.csl
bibliography:
- references.bib
- knitcitations.bib
---

```{r include=FALSE}

#Load main libraries at once
library(tidyverse)
library(dplyr)
library(data.table)

#To obtain the same template
# install.packages("remotes")
#remotes::install_github("Pakillo/template")

```

```{r include=FALSE}
#Load data
data <- read.table("../Data/iberian_bees.csv.gz",  header=T, quote="\"", sep=",")
#Delete first col that are rownames
data <- data %>% select(-X)

#Load Iberianbees masterlist
master <- read.csv("../Data/Processing_iberian_bees_raw/Iberian_species_masterlist.csv", stringsAsFactors = FALSE)

#Load removed species from the database
removed <- read.csv("../Data/Processing_iberian_bees_raw/removed.csv")
master$accepted_name <- paste(master$Genus, master$Species, sep = " ")
geo <- length(which(!is.na(data$Latitude) & !is.na(data$Longitude)))
flw <- length(which(!is.na(data$Flowers.visited)))
```

@@

```{r Tabla1, results='asis', echo=FALSE, cache=FALSE}
library("knitr")
metadata <- read.csv("../Data/Metadata/attributes.csv")
kable(metadata[,-1])
```

@@

```{r Fig2, echo=FALSE, fig.cap="Figura 2. Segundo ejemplo", message=FALSE, warning=FALSE, cache=FALSE}
library(sf)
library(raster)
library(rasterVis)
#code courtesy of Paco Rodriguez.
crs.geo <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
crs.laea <- "+proj=laea +lat_0=52 +lon_0=10 +x_0=4321000 +y_0=3210000 +ellps=GRS80 +units=m +no_defs"

#Convert coordinates to numeric 
data$Longitude <- as.numeric(data$Longitude)
data$Latitude <- as.numeric(data$Latitude)
#Filter out missing coords to work with them
dat <- subset(data, !is.na(Latitude) & !is.na(Longitude))

datos.sf <- st_as_sf(dat,coords = c("Longitude", "Latitude"),
crs = crs.geo)

@@

ras.counts <- rasterize(datos.laea[, "accepted_name"], ras,
               field = "accepted_name", fun = "count")
#str(datos.laea)
datos.laea$total <- datos.laea$Male + datos.laea$Female + datos.laea$Worker + datos.laea$Not.specified 
ras.counts2 <- rasterize(datos.laea[, "total"], ras,
               field = "total", fun = "sum")
ras2 <- raster(datos.laea, resolution = 30000)
ras.richness <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {length(unique(na.omit(x)))} )

library(vegan)
ras.chao <- rasterize(datos.laea[, "accepted_name"], ras2,
               field = "accepted_name", fun = function(x, ...)
                 {estimateR(t(as.matrix(x = table(x), nrow = 1, ncol = x)))[2,1]} ) #2 chao, 4 ACE #TO DO THIS WELL, we need to recover abundances.
